---
title: "homework1"
author: liam albert
output: github_document
date: "2026-09-25"
---

## Loading packages
```{r message= F}
library(tidyverse)

data("penguins", package = "palmerpenguins")


```




## Question 1

**Data Set Description **

* There are 3 species of penguins listed; Adelie, Gentoo, and Chinstrap.

* The important variables are: species, island, bill_length_mm, bill_depth,mm, flipper_length_mm, body_mass_g, sex, and year.

* The penguins dataset contains `r nrow(penguins)` observations and `r ncol(penguins)` variables.

* The mean flipper legnth is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.



```{r}
names(penguins)

head(penguins)
```

**Building a Scatter Plot**

```{r}
ggplot(data = penguins, 
      aes(x = bill_length_mm, y = flipper_length_mm, color = species) ) + geom_point()

ggsave("penguin_plot.png")

```

## Question 2

Create a data frame comprised of:

* a random sample of size 10 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 10
* a factor vector of length 10, with 3 different factor “levels”

**Building a data frame:**
```{r}
df = tibble(
random_sample = rnorm(10),

logical_vec = random_sample > 0,

character_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),

factor_vec = factor(c("A", "B", "C","A", "B", "C","A", "B", "C","A"))
)

df


```

**Means of each variable:**

* The mean of random_sample is `r mean(pull(df,random_sample))`

* The mean of logical_vec is `r mean(pull(df,logical_vec))`



```{r}
mean(pull(df,random_sample))

mean(pull(df,logical_vec))

mean(pull(df,character_vec))

mean(pull(df,factor_vec))
```

**Variable conversion**
```{r results = 'hide'}

as.numeric(pull(df,logical_vec))

as.numeric(pull(df,character_vec))

as.numeric(pull(df,factor_vec))
```

### Explanation

These `mean` command works for these values because random_sample is already numeric, and R treats logical_vec as a binary of 1=true and 0=false. When we run the same command on character variables, we are returned with an `NA` and warning message. By applying `as.numeric` to the factor_vec we can convert the values to represent the 3 levels. This shows that R can calculate a mean when values are numeric and although R can convert factor vectors to numeric type, the values are not meaningful, so taking their means would not have any interpretation.

